Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-10-14 files/Chapter5/2D_FEM.tex
chapter 5 fixes

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, the Finite Element Method (FEM) is applied to the two- and three-dimensional cantilever beams as well as a Reissner-Mindlin cantilever plate. The aim of this chapter is to obtain a numerical formula to calculate the eigenvalues and eigenvectors for these models. This chapter is preparation for Chapter 6.\\

The outline of the chapter is as follows.
\begin{itemize}
	\item[] Section \ref{sec:FEM:2D} FEM for a two-dimensional cantilever body.
	\item[] Section \ref{sec:FEM:3D} FEM for a three-dimensional cantilever body. 
	\item[] Section \ref{sec:FEM:Plate} FEM for A Reissner-Mindlin cantilever plate model. 
\end{itemize}

\section{Two-dimensional cantilever body} \label{sec:FEM:2D}
Consider a two-dimensional cantilever elastic body, with a rectangular cross-section.

\subsubsection{Reference configuration for rectangular cross-section}
Let $\{e_1,e_2\}$ be a right-handed orthonormal basis for $R^2$. Denote the elastic body by $\Omega \in R_2$, with $(0,0)$ as the reference point. For a rectangular cross-section, $\Omega$ can be described as
\begin{eqnarray*}
	\Omega = \left \{ x \in E_2 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} \right \},
\end{eqnarray*} where $\partial \Omega$ denotes the boundary of $\Omega$. The boundary $\partial \Omega$ can be divided into the four distinct lines as follows:

\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Sigma_0:& \quad x_1 &= 0\\
		\Sigma_1:& \quad x_1 &= 1
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Gamma_1:& \quad x_2 &= -{h}/{2}\\
		\Gamma_2:& \quad x_2 &= {h}/{2}
	\end{eqnarray*}
\end{minipage}

\subsubsection{Cantilever elastic body}
Consider a two-dimensional elastic body clamped rigidly to a surface at $x_1 = 0$. The body is free-hanging on the other boundaries This is Problem 2D-1 in section \ref{ssec:2D_Model:ModelProblem}.
\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}

@@

@@

In section \ref{ssec:2D_Model:VariationalForm} the variational problem for the two dimensional cantilever model is defined by Problem 2D-1V. For convenience, some results from section \ref{ssec:2D_Model:VariationalForm} are repeated here.

\subsubsection{Problem 2D-1V}\label{sssec:2D_Model:Problem2D1VChp5}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{eq:2D_Model:Problem2D1VEqChp5}
\end{align}
for all $\phi \in T(\Omega)$.\\

With the test function space 
\begin{eqnarray*}
	T(\Omega) & = & \left\{ \phi \in C^1(\bar{\Omega}) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}.
\end{eqnarray*}
The bilinear forms and integral function defined by
\begin{eqnarray}
	b(u,\phi) & = & \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dA, \label{eq:2D_Model:BilinearChp5}\\
	c(\partial_t^2 u,\phi) & = & \int_\Omega~ (\partial^2_t u) \cdot \phi~dA, \label{eq:2D_Model:Bilinear_cChp5}\\
	(f,g) &=& \int_{\Omega} f\cdot g \ dA, \label{eq:2D_Model:Bilinear_intChp5}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1-\nu^2)}$.\\

Using the definition of the reference configuration, the constitutive equations and the bilinear form $b$ can be rewritten as follows.

@@

\subsection{Galerkin approximation}\label{2d_FEM_G}
Divide the reference configuration $\Omega$ into a rectangular grid of elements, such that there are $n = n_1\times n_2$ nodes.\\

Define a set of $n$-dimensional linear independent basis functions. For the two dimensional model, the basis functions can be defined by the set $$\left\{\langle\phi_1, 0\rangle, \langle\phi_2, 0\rangle,...,\langle\phi_{n}, 0 \rangle,\langle 0,\phi_1\rangle,\langle 0 ,\phi_2\rangle,...,\langle 0,\phi_{n}\rangle \right\}.$$ For this dissertation, we use bi-cubic functions $\phi_i$, however simpler bi-linear functions can also be used. For easier notation, denote each of these basis functions by $\delta_i$.\\

The functions in this set that satisfies the conditions of the test function space $T(\Omega)$ are called the admissible basis functions. The admissible basis functions for this model can be renumbered and expressed as the set $\left \{\delta_1, \delta_2,..., \delta_k \right\}$ for some $k \leq 2n$.\\

@@

Problem 2D-1V can now be rewritten into a Galerkin Approximation, denoted by Problem 2D-1G.

\subsubsection{Problem 2D-1G}
Find a function $u^h$ such that for all $t>0$, $u^h \in S^h$ and
\begin{eqnarray*}
	(u^h, \phi_i) & = & -b(u^h,\phi_i) + (Q^I, \phi_i)
\end{eqnarray*} for $i = 1,2,...,k$.\\

\textbf{Remarks:}
\begin{itemize}
	\item[] $Q^I$ is scalar vector obtained after interpolating the function $Q$ over the rectangular grid $\Omega$. i.e. $Q^I_{i,j} = Q(x_i,x_j)$ for $i = 1,2,...,n_1$ and $j = 1,2,...,n_2$.
\end{itemize}

@@

Consider the following standard Finite Element Method matrices
\subsubsection{FEM matrices}
\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\mathbf{M}_{j,i} & = & \int_{\Omega} \phi_i \phi_j ~dA \\
		\mathbf{{K}_{11}}_{j,i} & = & \int_{\Omega} \partial_1\phi_i \partial_1\phi_j~dA\\
		\mathbf{{K}_{22}}_{j,i} & = & \int_{\Omega} \partial_2\phi_i \partial_2\phi_j~dA
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\mathbf{{K}_{12}}_{j,i} & = & \int_{\Omega} \partial_2\phi_i \partial_1\phi_j~dA\\
		\mathbf{{K}_{21}}_{j,i} & = & \int_{\Omega} \partial_1\phi_i \partial_2\phi_j~dA
	\end{eqnarray*}
\end{minipage}
for $i,j = 1,2,...,k$.\\

And 
\begin{eqnarray*}
	{M_{F}}_{j,i} = \int_{\Omega}  \phi_i \phi_j~dA
\end{eqnarray*} for $i = 1,2,...,k$ and for $j =1,2,...,2n$.\\

@@

The matrices ${O}$ and ${O_f}$ are the zero matrices of the same size as $\mathbf{M}$ and ${M_f}$ respectively.\\

@@

\subsubsection{Remark}
This form of $\bar{u}$ is determined by the Hermite cubic basis functions.

@@

Now the eigenvalue problem can be solve using a finite difference method.

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, the Finite Element Method (FEM) is applied to the models of this dissertation, presented in chapter 1. The goal of this chapter is to obtain numerical formulas that can be used to calculate the eigenvalues and eigenvectors for the models. This is in preperation of chapter 6, where the eigenvalues and eigenvectors of the models are compared.

In chapter 4, a method was discussed to obtain the eigenvalues and eigenvectors for the Timoshenko beam model. Therefore it is not neccesary to apply the Finite Element Method on the Timoshenko model to obtain the eigenvalues and eigenvectors. The Finite Element Mehtod is applied to the two-dimensional and three-dimensional models.

The outline of the chapter is as follows.
\begin{itemize}
	\item[] Section \ref{sec:FEM:2D} FEM for a cantilever two-dimensional elastic body. This is problem 2D-1 in section \ref{ssec:2D_Model:ModelProblem}.
	\item[] Section \ref{sec:FEM:3D} FEM for a cantilever three-dimensional elastic body. This is problem 3D-1 in section \ref{ssec:3D_Model:ModelProblems}.
	\item[] Section \ref{sec:FEM:Plate} FEM for a cantilever Reissner-Mindlin plate. This is problem P-1 in section \ref{ssec:P_Model:ModelProblems}.
\end{itemize}

\section{A cantilever two-dimensional body} \label{sec:FEM:2D}
Consider a cantilever two-dimensional elastic body, with a rectangular cross-section.

\subsubsection{Reference configuration for rectangular cross-section}\label{sssec:2D_Model:RefConf}
Let $\{e_1,e_2\}$ be a right-handed orthonormal basis for $R^2$. Denote the elastic body by $\Omega \in R_2$, with $(0,0)$ as the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as,
\begin{eqnarray*}
	\Omega = \left \{ x \in E_2 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} \right \},
\end{eqnarray*} where $\partial \Omega$ denotes the boundary of $\Omega$. The boundary $\partial \Omega$ can be divided into the four distinct lines (or edges) as follows:

\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Sigma:& \quad x_1 &= 0\\
		\Gamma_3:& \quad x_1 &= 1
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Gamma_1:& \quad x_2 &= -{h}/{2}\\
		\Gamma_2:& \quad x_2 &= {h}/{2}
	\end{eqnarray*}
\end{minipage}

In this configuration, the body is clamped rigidly to the surface at $x_1 = 0$ denoted as $\Sigma$ and the body is free-hanging on the other boundaries denoted by $\Gamma$.

\subsubsection{Cantilever elastic body}
Consider a two-dimensional elastic body clamped rigidly to a surface at $x_1 = 0$. The body is free-hanging on the other boundaries. Recall this is Problem 2D-1 in section \ref{ssec:2D_Model:ModelProblem}.
\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}

@@

		% Add Σ at x_1 = 0
		\node at (-1.4,0) {$\Sigma$};

		% Add Γ₁ at x_2 = -h/2
		\node at (3.5,-1.2) {$\Gamma_1$};

		% Add Γ₂ at x_2 = h/2
		\node at (3.5,0.9) {$\Gamma_2$};

		% Add Γ₃ at x_1 = 1
		\node at (8,0) {$\Gamma_3$};

@@

In section \ref{ssec:2D_Model:VariationalForm} the variational problem for the two dimensional cantilever model is derived and refered to as Problem 2D-1V. Some of the results from section \ref{sec:2D_Model} are repeated here for convenience. Using the reference configuration, these results are rewritten from a general form to a model specific form.

\subsubsection{Problem 2D-1V}\label{sssec:2D_Model:Problem2D1VChp5}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{eq:2D_Model:Problem2D1VEqChp5}
\end{align}
for all $\phi \in T(\Omega)$. With the test function space 
\begin{eqnarray*}
	T(\Omega) & = & \left\{ \phi \in C^1(\bar{\Omega}) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}.
\end{eqnarray*}

The bilinear forms and integral function is defined by
\begin{eqnarray}
	b(u,\phi) & = & \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dA, \label{eq:2D_Model:BilinearChp5}\\
	c(\partial_t^2 u,\phi) & = & \int_\Omega~ (\partial^2_t u) \cdot \phi~dA, \label{eq:2D_Model:Bilinear_cChp5}\\
	(f,g) &=& \int_{\Omega} f\cdot g \ dA, \label{eq:2D_Model:Bilinear_intChp5}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1-\nu^2)}$.

Using the reference configuration, the constitutive equations and the bilinear form $b$ can be rewritten into a model specific form.

@@

See chapter 2 for the existence theory.

\subsection{Galerkin approximation}\label{2d_FEM_G}
To be able to apply the Finite Element Method to Problem 2D-1V, the body $\Omega$ needs to be discretised. This is done by dividing the body $\Omega$ into discrete shapes, called elements. There are various types of shapes of elements that can be used. Since the body $\omega$ is has a rectangular cross-section, rectangular are the simplist elements to use.

Divide the reference configuration $\Omega$ into a grid of rectangular elements, such that there are $n = n_1\times n_2$ nodes.

Define a set of $n$-dimensional linear independent basis functions. For the two dimensional model, the basis functions can be defined by the set $$B = \left\{\langle\phi_1, 0\rangle, \langle\phi_2, 0\rangle,...,\langle\phi_{n}, 0 \rangle,\langle 0,\phi_1\rangle,\langle 0 ,\phi_2\rangle,...,\langle 0,\phi_{n}\rangle \right\}.$$

These functions are chosen as bi-cubic functions $\phi_i$. Simpler bi-linear functions can also be used, however the use of the bi-cubic basis functions results in faster convergence and also the benefit of obtaining the derivatives of the solution at the expense of more complexity.

The subset of basis functions $B$ that satisfies all the conditions of the test function space $T(\Omega)$ are called the admissible basis functions. Denote the admissible basis functions by $\delta_j$, where each $\delta_j$ is a unique element of $B$. The admissible basis functions can be numbered and expressed as the set $A = \left \{\delta_1, \delta_2,..., \delta_k \right\}$ for some $k \leq 2n$.\\

@@

Substitution of $u^h$ into Problem 2D-1V, results in the following Galerkin Approximation, denoted by Problem 2D-1G.

\subsubsection{Problem 2D-1G}
Find a function $u^h$ such that for all $t>0$, $u^h \in S^h$ and
\begin{eqnarray*}
	(u^h, \phi_i) & = & -b(u^h,\phi_i) + (Q^I, \phi_i)
\end{eqnarray*} for $i = 1,2,...,k$. $Q^I$ is scalar vector obtained after interpolating the function $Q$ over the rectangular grid $\Omega$. i.e. $Q^I_{i,j} = Q(x_i,x_j)$ for $i = 1,2,...,n_1$ and $j = 1,2,...,n_2$.

@@

Consider the following standard Finite Element Method matrices
\subsubsection{FEM matrices}
\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\mathbf{M}_{j,i} & = & \int_{\Omega} \phi_i \phi_j ~dA \\
		\mathbf{{K}_{11}}_{j,i} & = & \int_{\Omega} \partial_1\phi_i \partial_1\phi_j~dA\\
		\mathbf{{K}_{22}}_{j,i} & = & \int_{\Omega} \partial_2\phi_i \partial_2\phi_j~dA
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\mathbf{{K}_{12}}_{j,i} & = & \int_{\Omega} \partial_2\phi_i \partial_1\phi_j~dA\\
		\mathbf{{K}_{21}}_{j,i} & = & \int_{\Omega} \partial_1\phi_i \partial_2\phi_j~dA
	\end{eqnarray*}
\end{minipage}
for $i,j = 1,2,...,k$.

And 
\begin{eqnarray*}
	{M_{F}}_{j,i} = \int_{\Omega}  \phi_i \phi_j~dA
\end{eqnarray*}

for $i = 1,2,...,k$ and for $j =1,2,...,2n$.

@@

The matrices ${O}$ and ${O_f}$ are the zero matrices of the same size as $\mathbf{M}$ and ${M_f}$ respectively.

@@

\textbf{Remark} This form of $\bar{u}$ is determined by the use of the bi-cubic basis functions.

@@

### 2023-10-14 files/Chapter5/3D_FEM.tex
cleanup chapter 6
second part chapter 6
chapter 5 fixes

\section{Three-dimensional elastic model}\label{sec:FEM:3D}
Consider a three-dimensional cantilever elastic body, with a rectangular cross-section.

\subsubsection{Reference configuration for rectangular cross-section}
Let $\{e_1,e_2,e_3\}$ be a right-handed orthonormal basis for $R^3$. Denote the elastic body as $\Omega \in R^3$ with $(0,0,0)$ the reference point. For a rectangular cross-section, $\Omega$ can be described as
\begin{eqnarray*}
	\Omega = \left\{ x \in R^3 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} , \ -\frac{b}{2} \leq x_3 \leq \frac{b}{2}\right \}
\end{eqnarray*}
Let $\partial \Omega$ denote the boundary of the body. Divide $\partial \Omega$ into the six distinct flat surfaces as follows:

\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Sigma_0:& \quad x_1 &= 0\\
		\Sigma_1:& \quad x_1 &= 1\\
		\Gamma_1:& \quad x_3 &= -{b}/{2} 
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Gamma_2:& \quad x_2 &= -{h}/{2}\\
		\Gamma_3:& \quad x_3 &= {b}/{2}\\
		\Gamma_4:& \quad x_2 &= {h}/{2} 
	\end{eqnarray*}
\end{minipage}

\subsubsection{Cantilever elastic body}
Consider a three-dimensional elastic body with rectangular cross-section, rigidly clamped to a surface at attached at the side $\Sigma_0$ and free-hanging at all the other sides. This is referred to as Problem 3D-1 in section \ref{ssec:3D_Model:ModelProblems}.
\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}

@@

In section \ref{ssec:3D_Model:VariationalForm}, the variational problem for the three-dimensional cantilever model is defined by Problem 3D-1V. For convenience, some results from section \ref{ssec:3D_Model:VariationalForm} are repeated here.

@@

Using the definition of the reference configuration, the constitutive equations and the bilinear form $b$ can be rewritten as follows.
\subsubsection*{Constitutive equations}
\begin{eqnarray*}
	\sigma_{11} & = &  \frac{1}{\gamma(1+\nu)}\partial_1 u_1 + \frac{\nu}{\gamma (1+\nu)(1-2\nu)}(\partial_1 u_1 +  \partial_2 u_2 + \partial_3 u_3)\label{3DB_3}\\
	\sigma_{22} & = &  \frac{1}{\gamma(1+\nu)}\partial_2 u_2 + \frac{\nu}{\gamma (1+\nu)(1-2\nu)}(\partial_1 u_1 +  \partial_2 u_2 + \partial_3 u_3)\label{3DB_4}\\
	\sigma_{33} & = &  \frac{1}{\gamma(1+\nu)}\partial_3 u_3 + \frac{\nu}{\gamma (1+\nu)(1-2\nu)}(\partial_1 u_1 +  \partial_2 u_2 + \partial_3 u_3)\label{3DB_5}\\
	\sigma_{23} & = &  \frac{1}{2\gamma(1+\nu)}(\partial_3 u_2 + \partial_2 u_3)\label{3DB_6}\\
	\sigma_{31} & = &  \frac{1}{2\gamma(1+\nu)}(\partial_3 u_1 + \partial_1 u_3)\label{3DB_7}\\
	\sigma_{12} & = &  \frac{1}{2\gamma(1+\nu)}(\partial_2 u_1 + \partial_1 u_2)\label{3DB_8}
\end{eqnarray*}

@@

In this form, the bilinear form $b$ is more useful for applications than the general form.

@@

\subsection{Galerkin approximation}
Divide the reference configuration $\Omega$ into a grid of rectangular prismatic elements, such that there are $n = n_1 \times n_2 \times n_3$ nodes. These rectangular prismatic elements are also known as brick shaped elements as described by \cite{Wu06}. This shape of element is a natural choice for a three-dimensional elastic body with a rectangular cross-section.\\

Define a set of $n$-dimensional linear independent basis functions. For the three-dimensional model, the basis functions can be defined by the set
\begin{eqnarray*}
 \{\langle\phi_1, 0 , 0\rangle, \langle\phi_2, 0, 0\rangle,...,\langle\phi_{n}, 0, 0 \rangle,\\
	\langle 0,\phi_1 ,0 \rangle,\langle 0 ,\phi_2,0\rangle,...,\langle 0,\phi_{n},0\rangle,\\
	\langle 0,0,\phi_1 \rangle,\langle 0,0,\phi_2\rangle,...,\langle 0,0,\phi_{n}\rangle \}.
\end{eqnarray*}

@@

The functions that satisfy the conditions of the test function space $T(\Omega)$ are called the admissible basis functions. The admissible basis functions for this model can be renumbered and expressed as the set $\left\{\delta_1, \delta_2, ..., \delta_k \right\}$ for some $k \leq 3n$.\\

@@

Problem 3D-1V can now be rewritten into a Galerkin Approximation, denoted by Problem 3D-1G.

\subsubsection{Problem 3D-1G}
Find a function $u^h$ such that for all $t>0$, $u^h \in S^h$ and
\begin{eqnarray*}
	(u^h, \phi_i) & = & -b(u^h,\phi_i) + (Q^I \cdot \phi_i)
\end{eqnarray*} for $i = 1,2,...,k$.\\

@@

\subsubsection{Problem 3D-1ODE}
Find a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	M\ddot{\bar{u}} & = & K\bar{u} + M_{f}Q^I. \label{3D_M}
\end{eqnarray} With $\bar{u}$ in the form $\bar{u} = \langle u, \partial_1 u, \partial_2 u, \partial_3 u, \partial_{11} u, \partial_{12}u, \partial_{13}u,\partial_{22}u,\partial_{23}u, \partial_{33}u,$\\$ \partial_{111}u, \partial_{112}u, \partial_{113}u, \partial_{121}u, \partial_{122}u, \partial_{123}u, \partial_{133}u,\partial_{223}u,\partial_{233}u,\partial_{333}u \rangle$. \textcolor{red}{beter manier om voor te stel?}

\subsection{Eigenvalue problem}
The equation \eqref{3D_M} is the same form as in section \ref{2dFEM_EP} for the two-dimensional elastic body. Therefore the derivation of the eigenvalue problem is identical. Denote the eigenvalue problem for Problem 3D-1 by Problem 3D-1E.

\subsubsection{Problem 3D-1E}
Find a real number $\lambda$ and a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	M\lambda{\bar{u}} & = & K\bar{u}.
\end{eqnarray}
\end{document}

>>>>

In this section, a similar approach is applied to the three-dimensional elastic body.

\section{A cantilever three-dimensional body}\label{sec:FEM:3D}
Consider a cantilever three-dimensional elastic body, with a rectangular cross-section.

\subsubsection{Reference configuration for rectangular cross-section}
Let $\{e_1,e_2,e_3\}$ be a right-handed orthonormal basis for $R^3$. Denote the elastic body as $\Omega \in R^3$ with $(0,0,0)$ the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as
\begin{eqnarray*}
	\Omega = \left\{ x \in R^3 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} , \ -\frac{b}{2} \leq x_3 \leq \frac{b}{2}\right \}
\end{eqnarray*}
Let $\partial \Omega$ denote the boundary of the body. Divide $\partial \Omega$ into the six distinct flat surfaces as follows:

\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Sigma:& \quad x_1 &= 0\\
		\Gamma_5:& \quad x_1 &= 1\\
		\Gamma_1:& \quad x_3 &= -{b}/{2} 
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Gamma_2:& \quad x_2 &= -{h}/{2}\\
		\Gamma_3:& \quad x_3 &= {b}/{2}\\
		\Gamma_4:& \quad x_2 &= {h}/{2} 
	\end{eqnarray*}
\end{minipage}

Using this notation, the boundary conditions are similar to the two-dimensional model in section \ref{sssec:2D_Model:RefConf}, where the body is clamped rigidly at $\Sigma$, and free-hanging on the other sides denoted by $\Gamma$.

\subsubsection{Cantilever elastic body}
Consider a three-dimensional elastic body with rectangular cross-section, rigidly clamped to a surface at attached at the side $\Sigma$ and free-hanging at all the other sides. This is Problem 3D-1 in section \ref{ssec:3D_Model:ModelProblems}.
\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}

@@

In section \ref{ssec:3D_Model:VariationalForm}, the variational problem for the three-dimensional cantilever model is defined by Problem 3D-1V. This general form can now be rewritten as in a model specific form using the reference configuration. For convenience, some of the results from section \ref{sec:3D_Mode} are repeated here.

@@

Using the definition of the reference configuration, the constitutive equations and the bilinear form $b$ can be rewritten in the following model specific forms: 
\subsubsection*{Constitutive equations}
\begin{eqnarray*}
	\sigma_{11} & = &  \frac{1}{\gamma(1+\nu)}\partial_1 u_1 + \frac{\nu}{\gamma (1+\nu)(1-2\nu)}(\partial_1 u_1 +  \partial_2 u_2 + \partial_3 u_3)\label{3DB_3}\\
	\sigma_{22} & = &  \frac{1}{\gamma(1+\nu)}\partial_2 u_2 + \frac{\nu}{\gamma (1+\nu)(1-2\nu)}(\partial_1 u_1 +  \partial_2 u_2 + \partial_3 u_3)\label{3DB_4}\\
	\sigma_{33} & = &  \frac{1}{\gamma(1+\nu)}\partial_3 u_3 + \frac{\nu}{\gamma (1+\nu)(1-2\nu)}(\partial_1 u_1 +  \partial_2 u_2 + \partial_3 u_3)\label{3DB_5}\\
	\sigma_{23} & = &  \frac{1}{2\gamma(1+\nu)}(\partial_3 u_2 + \partial_2 u_3)\label{3DB_6}\\
	\sigma_{31} & = &  \frac{1}{2\gamma(1+\nu)}(\partial_3 u_1 + \partial_1 u_3)\label{3DB_7}\\
	\sigma_{12} & = &  \frac{1}{2\gamma(1+\nu)}(\partial_2 u_1 + \partial_1 u_2)\label{3DB_8}
\end{eqnarray*}

@@

@@

See chapter 2 for the existence theory.

\subsection{Galerkin approximation}
As mentioned in the previous section, section \ref{2d_FEM_G}, the body $\Omega$ needs to be descritised. This is model, three-dimensional elements shapes are required. The elements used in this dissertation are rectangular prismatic elements. These elements are also known as brick shaped elements as described by \cite{Wu06}. This shape of element is a natural choice for a three-dimensional elastic body with a rectangular cross-section.\

Divide the reference configuration $\Omega$ into a grid of rectangular prismatic elements, such that there are $n = n_1 \times n_2 \times n_3$ nodes.

Define a set of $n$-dimensional linear independent basis functions. For the three-dimensional model, the basis functions can be defined by the set
\begin{eqnarray*}
 B = \{\langle\phi_1, 0 , 0\rangle, \langle\phi_2, 0, 0\rangle,...,\langle\phi_{n}, 0, 0 \rangle,\\
	\langle 0,\phi_1 ,0 \rangle,\langle 0 ,\phi_2,0\rangle,...,\langle 0,\phi_{n},0\rangle,\\
	\langle 0,0,\phi_1 \rangle,\langle 0,0,\phi_2\rangle,...,\langle 0,0,\phi_{n}\rangle \}.
\end{eqnarray*}

For the this three-dimensional model, tri-cubix basis functions are used. These basis functions are an extension of the Hermite cubic basis functions from the two-dimensional model. Although this is more complex than using tri-linear basis functions, as mentioned in section \ref{2d_FEM_G}, the tri-cubic basis functions ensure faster convergence and also the derivatives of the solutions.\\

@@

Similar to \ref{2d_FEM_G}, denote the admissible basis functions by $\delta_j$, where each $\delta_j$ is a unique element of $B$ satisfying all the conditions of the test function space $T(\Omega)$. The admissible basis functions are numbered and expressed as the set $A = \left\{\delta_1, \delta_2, ..., \delta_k \right\}$ for some $k \leq 3n$.\\

@@

Substituting $u^h$ into Problem 3D-1V, results in the following Galerkin approximation, denoted by Problem 3D-1G.

\subsubsection{Problem 3D-1G}
Find a function $u^h$ such that for all $t>0$, $u^h \in S^h$ and
\begin{eqnarray*}
	(u^h, \phi_i) & = & -b(u^h,\phi_i) + (Q^I \cdot \phi_i)
\end{eqnarray*} for $i = 1,2,...,k$. $Q^I$ is scalar vector obtained after interpolating the function $Q$ over the rectangular grid $\Omega$. i.e. $Q^I_{i,j, h} = Q(x_i,x_j, x_h)$ for $i = 1,2,...,n_1$, $j = 1,2,...,n_2$ and $h = 1,2,...,n_3$.

@@

\subsubsection{Problem 3D-1ODE}
Find a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	M\ddot{\bar{u}} & = & K\bar{u} + M_{f}Q^I. \label{3D_M}
\end{eqnarray} With $\bar{u}$ in the form $\bar{u} = \partial^i_1u \partial^j_2u \partial^k_3u$ for $i,j,k = 0,1,2,3$.

\subsection{Eigenvalue problem}
The derivation and form of the eigenvalue problem for the three-dimensional elastic body is similar to the two-dimensional model given in section \ref{2dFEM_EP}.

\subsubsection{Problem 3D-1E}\label{3dFEM_EP}
Find a real number $\lambda$ and a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	M\lambda{\bar{u}} & = & K\bar{u}.
\end{eqnarray}

\end{document}
